---
title: "Convergence Statistics"
author: "Karl Hornlund"
date: "September 2, 2018"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The following is an analysis of confidence interval sizes for Bernoulli Trials. In particular: 

*How does the width of a 95% confidence interval narrow as the number of trials increases?*

The goal is to inform decisions choosing a number of trials $N$, such that $N$ is not *unnecessarily large*, while large enough to give appropriate *confidence* in estimates resulting from the trials.


```{r funcs, echo = FALSE, message=FALSE}
library(tidyverse)
library(ggthemes)
library(plotly)

probs <- seq(0.50, 0.99, 0.01)
N     <- seq(50, 1000, 1)
runs  <- seq(200, 800, 100)

width_func <- function(p, N) {
  std_dev <- sqrt(p * (1-p))
  z <- qnorm(.975) # 95% CI
  width <- 2 * z * std_dev/sqrt(N)
  return(width)
}

df <- expand.grid(N, probs)
names(df) <- c("N", "P")

df <- df %>%
  mutate(width = width_func(.$P, .$N))
```

## Results

The width of the confidence interval is computed as follows:

$w = f(p, N) = 2z \: \frac{\sqrt{p(1-p)}}{\sqrt{N}} \quad , \quad z = 1.96$

### General Trends 

The following shows the expected width of a 95% Confidence Interval, given the probability of success $p$, and the number of trials performed $N$.

```{r plot01, echo=FALSE}
gg1 <- df %>%
  filter(N %in% runs) %>%
  ggplot(aes(x = P, y = width, color = factor(N))) +
  geom_line() +
  scale_color_gdocs() +
  theme_gdocs() +
  labs(
    title = "Confidence Interval Widths for Bernoulli Trials",
    x = "Probability of Bernoulli Trial Success (p)",
    y = "E[ Width of 95% Confidence Interval ]",
    color = "Trials (N)"
  )

ggplotly(gg1)
```

### Intersection with $w = 0.05$

The following shows the number of trials ($N$) required to narrow the Confidence Interval width to below $0.05$, for a given probability $p$. In other words, it shows how the lines in the plot above intersect with $y = 0.05$.

```{r plot02, echo=FALSE}
min.df <- df %>%
  filter(width <= 0.05) %>%
  group_by(P) %>%
  summarise(min_runs = min(N))

gg2 <- min.df %>%
  ggplot(aes(x = P, y = min_runs)) +
  geom_line() +
  scale_color_gdocs() +
  theme_gdocs() +
  scale_x_continuous(breaks = seq(0.8, 1, 0.02)) +
  labs(
    title = "Trials required for 0.05 Confidence Interval Width",
    x = "Probability of Bernoulli Trial Success (p)",
    y = "E[ Trials ]"
  )

ggplotly(gg2)
```
